---
title: "AWS RDS MySQL Connection and Query"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This R Markdown document demonstrates how to connect to an AWS RDS MySQL database using R, and perform basic database operations.

## Loading Required Packages

First, we'll load the necessary packages. If they're not installed, we'll install them.

```{r load_packages}
if (!requireNamespace("DBI", quietly = TRUE)) install.packages("DBI")
if (!requireNamespace("RMySQL", quietly = TRUE)) install.packages("RMySQL")

library(DBI)
library(RMySQL)
```

## Setting Up Connection Parameters

We'll set up our connection parameters. For security, we're using environment variables for sensitive information.

```{r setup_params}
db_host <- "nepal-dashboard-db.c5nzpjrxfrcj.us-east-1.rds.amazonaws.com"
db_name <- "fsp_data"
db_user <- "admin"
db_password <- Sys.getenv("DB_PASSWORD")  # Make sure to set this environment variable
```

## Connecting to the Database

Now, let's create a function to connect to our database.

```{r connect_function}
connect_to_db <- function() {
  dbConnect(MySQL(),
            host = db_host,
            dbname = db_name,
            user = db_user,
            password = db_password)
}
```

## Database Operations

Here, we'll perform some basic database operations.

```{r db_operations}
tryCatch({
  # Establish connection
  con <- connect_to_db()
  cat("Connected to the database successfully.\n")
  
  # List all tables
  tables <- dbListTables(con)
  cat("Tables in the database:\n")
  print(tables)
  
  # Example query - modify as needed
  query <- "SELECT * FROM your_table_name LIMIT 5"
  result <- dbGetQuery(con, query)
  cat("\nSample data from your_table_name:\n")
  print(result)
  
  # You can add more database operations here
  
}, error = function(e) {
  cat("An error occurred:\n", conditionMessage(e), "\n")
}, finally = {
  # Close the connection if it's open
  if (exists("con") && dbIsValid(con)) {
    dbDisconnect(con)
    cat("Database connection closed.\n")
  }
})
```

## Conclusion

This document demonstrated how to connect to an AWS RDS MySQL database, list tables, and run a basic query. You can expand on this to perform more complex database operations as needed.

Remember to keep your database credentials secure and never share them publicly. It's a good practice to use environment variables for sensitive information.
