---
title: "WFP Database"
author: "Nanda Kumar Maharjan"
date: "2024-06-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Basic Packages for R studio
```{r warning=FALSE, message=FALSE}
rm(list = ls())
library(pacman)
p_load(tidyverse, dplyr, readxl, flextable, lubridate, here)

```



#Necessary library to connect R to SQL database 

```{r warning= FALSE, message=F}
library(odbc) # Contains drivers to connect a database
library(DBI)  # Contains function to interact with database
library(RSQLite)
library(dbplyr)
library(dplyr)

#portaldb <- dbConnect()
```

#Example to connect to SQL
```{r}
#con <- DBI::dbConnect(drv = odbc::odbc(),
#                      Driver = "driver_name",
#                      Server = "server_url",
#                      Database = "database_name",
#                      user = "user", #optional
#                      password = "password") #optional
```

```{r}
con <- DBI::dbConnect(drv = odbc::odbc(),
                      Driver = "Amazon_Redshift",
                      Server = "nepal-dashboard-db.c5nzpjrxfrcj.us-east-1.rds.amazonaws.com",
                      Database = "fsp_data",
                      user = "admin", #optional
                      password = "Q3aikjweEVYoox6yyZhi") #optional
```

#WFP Database from Python 
# Download the CSV file

```{r}
csv_data <- read.csv("https://data.humdata.org/dataset/0e8b1b84-8e55-4aec-bc45-0c7d2c92d32e/resource/10bae3d7-82bf-4caf-a216-5fb4a8ff6e43/download/wfp_food_prices_npl.csv")

data <- csv_data

rm(csv_data)

data %>% filter(date == "2023-05-15" | date == "2024-05-15" |date == "2024-04-15")
```

# Print the first few rows of the CSV data
```{r}
head(csv_data)
```
# REMOVE THE FIRST ROW, SINCE THERE IS UNNECESSARY INFORMATION

```{r}
csv_data <- csv_data[-1,]
csv_data
```
# ADD ISO3CODE SINCE THERE IS NO COUNTRYCODE ITSELF FROM THE CSV FILE

```{r}
csv_data <- csv_data %>% mutate(ISO3Code = 'NPL')
df <- csv_data
```


```{r}
tables <- c('WFP_PRICE_DATA', 'WFP_PRICE_DATA_PROCESSED')

truncate_existing_tables <- function(conn, tables) {
  for (table in tables) {
    dbExecute(conn, paste0("TRUNCATE TABLE ", table))
  }
}


insert_dataframe_to_db <- function(conn, df, table_name) {
  dbWriteTable(conn, table_name, df, append = TRUE, row.names = FALSE)
}

conn <- dbConnect(RSQLite::SQLite(), dbname = 'your_database.sqlite')

truncate_existing_tables(conn, tables)

insert_dataframe_to_db(conn, df, 'WFP_PRICE_DATA')
dbDisconnect(conn)
```

## Print the number of rows
```{r}
cat("Number of rows:", nrow(csv_data), "\n")
```
# Print the data type of the 'price' column
```{r}
cat("Data type of 'price':", class(csv_data$price), "\n")
```
# Filter the dataframe

```{r}
csv_data <- csv_data %>% filter(price != 0.0 & priceflag == "actual")
```
# Print the number of rows after filtering

```{r}
cat("Number of rows after filtering:", nrow(df), "\n")

nrow(csv_data)
```
#Calculating the annual percentage change for seasonality
# Copy the dataframe
```{r}
percentage_change_df <- df
```
# Convert 'date' to datetime format if it's not already
```{r}
percentage_change_df$date <- as.Date(percentage_change_df$date)
```

# Extract the year from the 'date' column and create a new column called 'year'
```{r}
percentage_change_df <- percentage_change_df %>%
  mutate(year = year(date))
```
# Create a column "month" to match with CPI data
```{r}
percentage_change_df <- percentage_change_df %>%
  mutate(month = month(date))

percentage_change_df %>% select(month)
```
# Convert 'price' to numeric value format if it's not already
```{r}
percentage_change_df <- percentage_change_df %>%
  mutate(price = as.numeric(price))
```

# Group by year, commodity, and market to calculate the mean price

```{r}
annual_prices <- percentage_change_df %>%
  group_by(year, commodity, market) %>%
  summarise(annual_avg_price = round(mean(price, na.rm = TRUE),2)) %>% ungroup()

annual_prices %>% flextable()
```
# View the resulting dataframe

```{r}
print(annual_prices)
```
# Merge the annual_prices dataframe back into the main percentage_change_df dataframe

```{r}
percentage_change_df <- percentage_change_df %>%
  left_join(annual_prices, by = c("year", "commodity", "market"))

percentage_change_df
```
# Calculate the percentage change between the actual price and the annual average price
```{r}
percentage_change_df <- percentage_change_df %>%
  mutate(percentage_change = ((price - annual_avg_price)/annual_avg_price * 100))

#To view percentage change
percentage_change_df %>% select(percentage_change)
```

# Define the function to insert the dataframe into the database
```{r}
insert_dataframe_to_db <- function(conn, df, table_name) {
  dbWriteTable(conn, table_name, df, append = TRUE, row.names = FALSE)
}
```

# Example usage
# Assuming you have a database connection already established
```{r}
conn <- dbConnect(RSQLite::SQLite(), dbname = 'your_database.sqlite')
```
# Insert the updated dataframe into the database

```{r}
insert_dataframe_to_db(conn, percentage_change_df, 'WFP_PRICE_DATA_PROCESSED')
```

# Close the connection
```{r}
dbDisconnect(conn)
```

#Calculating the real price, nominal index and real index for food prices tab

# Create a connection to the database
```{r}
conn <- dbConnect(RSQLite::SQLite(), dbname = 'your_database.sqlite')
```
# Define the SQL query
```{r}

query <- "SELECT value, CPI_month, CPI_year FROM NEPAL_DASHBOARD_CPI"
```

# Execute the query and fetch the results

```{r}
result <- dbGetQuery(conn, query)
```

# Convert the result to a dataframe
```{r}
cpi_df <- read_csv(here("data/CPI.csv"))



cpi_df <- as.data.frame(result)

```

# Print the column names of the dataframe
```{r}
print(colnames(cpi_df))
```
# Close the connection

```{r}
dbDisconnect(conn)
```

# Merge percentage_change_df with cpi_df on year and month

```{r}
percentage_change_CPI_df <- percentage_change_df %>%
  left_join(cpi_df, by = c("year" = "CPI_year", "month" = "CPI_month"))
```
# Drop unnecessary columns (if any)

```{r}
percentage_change_CPI_df <- percentage_change_CPI_df %>%
  select(-ends_with("_to_drop"))

percentage_change_CPI_df
```
# Change the column name of 'value' to 'CPI'

```{r}

percentage_change_CPI_df <- percentage_change_CPI_df %>%
  rename(value = CPI)

```

# Define the function to insert the dataframe into the database
```{r}
insert_dataframe_to_db <- function(conn, df, table_name) {
  dbWriteTable(conn, table_name, df, append = TRUE, row.names = FALSE)
}
```

# Example usage
# Assuming you have a database connection already established

```{r}

conn <- dbConnect(RSQLite::SQLite(), dbname = 'your_database.sqlite')

```


# Insert the updated dataframe into the database

```{r}
insert_dataframe_to_db(conn, percentage_change_CPI_df, 'WFP_PRICE_DATA_PROCESSED')
```

# Close the connection
```{r}
dbDisconnect(conn)
```

# Filter to get the real prices for January, February, and March of 2020
```{r}

percentage_change_CPI_real_df = percentage_change_CPI_df

jan_feb_mar_2020_real_prices <- percentage_change_CPI_real_df %>%
  filter(year == 2020 & month %in% c(1, 2, 3))

jan_feb_mar_2020_real_prices
```


# Filter to get missing real prices outside January, February, and March of 2020

```{r}
missing_real_prices <- percentage_change_CPI_real_df %>%
  filter(!(year == 2020 & month %in% c(1, 2, 3)))

missing_real_prices
```

# Further filter to get data points for the year 2020 and after
```{r}
missing_real_prices_2020_after <- missing_real_prices %>%
  filter(year >= 2020)
```

# Sort by 'admin1', 'admin2', 'market', and 'commodity'
```{r}
missing_real_prices_2020_after_sort <- missing_real_prices_2020_after %>%
  arrange(admin1, admin2, market, commodity)
```

# Reset the index (row names) in R
```{r}
missing_real_prices_2020_after_sort <- missing_real_prices_2020_after_sort %>%
  mutate(row_number = row_number()) %>%
  select(-row_number)
```

# View the result
```{r}
print(missing_real_prices_2020_after_sort)
```

# Assuming 'percentage_change_CPI_real_df' is already loaded

# Get the earliest values for each series

```{r}
earliest_real_values <- missing_real_prices_2020_after_sort %>%
  group_by(admin1, admin2, market, commodity) %>%
  slice(1) %>%
  ungroup() %>%
  rename(real_price_reference = real_price)
```
# Insert the resulting dataframe into the database

```{r}
conn <- dbConnect(RSQLite::SQLite(), dbname = 'your_database.sqlite')
insert_dataframe_to_db(conn, earliest_real_values, 'WFP_PRICE_DATA_PROCESSED_MISSING_REAL_REFERENCE')
```

# Calculate the average 'real_price' for each group ('commodity' and 'market')

```{r}
avg_real_prices_grouped <- jan_feb_mar_2020_real_prices %>%
  group_by(admin1, admin2, market, commodity) %>%
  summarise(real_price_reference = mean(real_price, na.rm = TRUE)) %>%
  ungroup()
```

# Merge this 'real_reference_price' from 2020 Jan back into the main dataframe
```{r}
percentage_change_CPI_real_index_df <- percentage_change_CPI_real_df %>%
  left_join(avg_real_prices_grouped, by = c('admin1', 'admin2', 'commodity', 'market'))
```


# Drop unnecessary columns (if any)
```{r}
percentage_change_CPI_real_index_df <- percentage_change_CPI_real_index_df %>%
  select(-ends_with("_to_drop"))
```

# Insert the updated dataframe into the database

```{r}
insert_dataframe_to_db(conn, percentage_change_CPI_real_index_df, 'WFP_PRICE_DATA_PROCESSED_REAL_REFERENCE1')

```

# Merge the earliest real values back into the main dataframe

```{r}
percentage_change_CPI_real_index_early_df <- percentage_change_CPI_real_index_df %>%
  left_join(
    earliest_real_values %>% select(admin1, admin2, commodity, market, real_price_reference),
    by = c('admin1', 'admin2', 'commodity', 'market'),
    suffix = c("", "_to_drop")
  )
```

# Drop unnecessary columns (if any)
```{r}
percentage_change_CPI_real_index_early_df <- percentage_change_CPI_real_index_early_df %>%
  select(-ends_with("_to_drop"))
```


# Insert the final dataframe into the database
```{r}
insert_dataframe_to_db(conn, percentage_change_CPI_real_index_early_df, 'WFP_PRICE_DATA_PROCESSED_REAL_REFERENCE_EARLY')
```

# Close the connection

```{r}
dbDisconnect(conn)
```

# Assuming 'percentage_change_CPI_real_index_early_df' and 'avg_real_prices_grouped' are already loaded

# Check if 'real_price_reference' is NA in percentage_change_CPI_real_index_early_df

```{r}
mask <- is.na(percentage_change_CPI_real_index_early_df$real_price_reference)
```


# Fill missing values in 'real_price_reference' from 'real_price_reference_to_drop'
```{r}
percentage_change_CPI_real_index_early_df$real_price_reference[mask] <- percentage_change_CPI_real_index_early_df$real_price_reference_to_drop[mask]
```


# Drop unnecessary columns (if any)
```{r}
percentage_change_CPI_real_index_early_df <- percentage_change_CPI_real_index_early_df %>%
  select(-ends_with("_to_drop"))
```


# Insert the updated dataframe into the database
```{r}
insert_dataframe_to_db(conn, percentage_change_CPI_real_index_early_df, 'WFP_PRICE_DATA_PROCESSED_REAL_REFERENCE2')
```

# Assuming 'percentage_change_CPI_real_nominal_index_df', 'jan_feb_mar_2020_nominal_prices', 'missing_nominal_prices', 'missing_nominal_prices_2020_after_sort', 'earliest_nominal_values', 'avg_nominal_prices_grouped' are already loaded

# Get the earliest values for each series

```{r}

earliest_nominal_values <- missing_nominal_prices_2020_after_sort %>%
  group_by(admin1, admin2, market, commodity) %>%
  slice(1) %>%
  ungroup() %>%
  rename(nominal_price_reference = price)
```


# Insert the resulting dataframe into the database
```{r}
insert_dataframe_to_db(conn, earliest_nominal_values, 'WFP_PRICE_DATA_PROCESSED_MISSING_NOMINAL_REFERENCE')
```


# Calculate the average 'price' for January, February, and March of 2020
```{r}
avg_nominal_prices_grouped <- jan_feb_mar_2020_nominal_prices %>%
  group_by(admin1, admin2, market, commodity) %>%
  summarise(nominal_price_reference = mean(price, na.rm = TRUE)) %>%
  ungroup()
```


# Merge this 'nominal_price_reference' from 2020 Jan back into the main dataframe
```{r}
percentage_change_CPI_real_nominal_index_df <- percentage_change_CPI_real_nominal_df %>%
  left_join(avg_nominal_prices_grouped, by = c('admin1', 'admin2', 'commodity', 'market'))
```


# Drop unnecessary columns (if any)
```{r}
percentage_change_CPI_real_nominal_index_df <- percentage_change_CPI_real_nominal_index_df %>%
  select(-ends_with("_to_drop"))
```


# Insert the updated dataframe into the database
```{r}
insert_dataframe_to_db(conn, percentage_change_CPI_real_nominal_index_df, 'WFP_PRICE_DATA_PROCESSED_NOMINAL_REFERENCE1')
```


# Merge the earliest nominal values back into the main dataframe
```{r}
percentage_change_CPI_nominal_index_early_df <- percentage_change_CPI_real_nominal_index_df %>%
  left_join(
    earliest_nominal_values %>% select(admin1, admin2, commodity, market, nominal_price_reference),
    by = c('admin1', 'admin2', 'commodity', 'market'),
    suffix = c("", "_to_drop")
  )
```


# Check if 'nominal_price_reference' is NA in percentage_change_CPI_nominal_index_early_df
```{r}
mask <- is.na(percentage_change_CPI_nominal_index_early_df$nominal_price_reference)
```


# Fill missing values in 'nominal_price_reference' from 'nominal_price_reference_to_drop'
```{r}
percentage_change_CPI_nominal_index_early_df$nominal_price_reference[mask] <- percentage_change_CPI_nominal_index_early_df$nominal_price_reference_to_drop[mask]
```


# Drop unnecessary columns (if any)
```{r}
percentage_change_CPI_nominal_index_early_df <- percentage_change_CPI_nominal_index_early_df %>%
  select(-ends_with("_to_drop"))
```


# Insert the updated dataframe into the database
```{r}
insert_dataframe_to_db(conn, percentage_change_CPI_nominal_index_early_df, 'WFP_PRICE_DATA_PROCESSED_NOMINAL_REFERENCE2')
```


# Calculate the nominal_price_index
```{r}
percentage_change_CPI_nominal_index_early_df$nominal_price_index <- (
  (percentage_change_CPI_nominal_index_early_df$price / percentage_change_CPI_nominal_index_early_df$nominal_price_reference) * 100
)
```


# Calculate the real_price_index
```{r}
percentage_change_CPI_nominal_index_early_df$real_price_index <- (
  (percentage_change_CPI_nominal_index_early_df$real_price / percentage_change_CPI_nominal_index_early_df$real_price_reference) * 100
)
```


# Insert the updated dataframe into the database
```{r}
insert_dataframe_to_db(conn, percentage_change_CPI_nominal_index_early_df, 'WFP_PRICE_DATA_NEPAL_PROCESSED')
```






